---
title: "Functional Time Transformation Model"
author: "Rahul Ghosal, Marcos Matabuena and Sujit Ghosh"
date: "07/01/2024"
output:
  html_document:
    toc: true
    number_sections: true
---
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document presents an illustration of the Functional Time Transformation Model (FTTM) developed in Ghosal et al., 2024. 

# Example 1: Functional Cox Model
First we generate survival times following a functional proportional hazard model (Scenario A1 in paper)and fit the FTTM. 
```{r, tidy = FALSE,message=FALSE,warning=FALSE, fig.width=6, fig.height=5}
###########Generating data##############################
set.seed(123)
n<-500  
Z1<-rbinom(n,1,0.5)
Z2<-rnorm(n,0,1)
X<-cbind(rep(1,n),Z1,Z2)
###############################
m <- 101
S<- seq(0, 1, length = m)
rankX <- 10
Phi <- cbind(1/sqrt(m), poly(S, degree=rankX-1))
lambda <- rankX:1
Xi <- sapply(lambda, function(l)
  scale(rnorm(n, sd=2*sqrt(l)), scale=FALSE))
Xs <- Xi %*% t(Phi) 

beta<-c(log(0.2),-0.5,0.4)
beta_s<-function(s){1*cos (pi*s)}
betaS<-beta_s(S)
lfterm2<-as.numeric((1/m)*Xs%*%betaS) #xs=zs

ratesurv<-function(i) 
{exp(sum(beta*X[i,])+lfterm2[i])
}
ratelat<-unlist(lapply(1:n,ratesurv)) #
mu_c<-20   #censoring times mean
T_tr<-c() ##survival times true
C<-c() #censoring time
T_obs<-c() #observed survival time
for(i in 1:n)
{
  temp<-rexp(1,rate=ratelat[i])
  T_tr[i]<- temp #true survival time
  C[i]<-rexp(1,rate=(1/mu_c))
  T_obs[i]=min(T_tr[i],C[i])
}
delta<-as.numeric(T_tr<C)

mean(delta==1) #proportion uncensored
mean(delta==0) #proportion censored 
###KM estimates#######
library(survival)
fitKM <- survfit(Surv(T_obs,delta) ~ 1)###
plot(fitKM, xlab="Days", main = 'Kaplan Meyer Plot') 
```

Next we fit a functional Cox model for illustration, which is the oracle model in this case.

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE, tidy=FALSE}
newdf<-data.frame("time"=T_obs,"status"=delta,"X1"=X[,2],"X2"=X[,3])
Xmat<-X[,c(2:3)]
library(mgcv)
Smat<-matrix(S,n,m,byrow=TRUE)
Xs_sc<-Xs/m
newdf$Xs_sc<-Xs_sc
mod1_func <-gam(time~X1+X2+s(Smat,k=20,by=Xs_sc,bs="ps",m=2),
                family=cox.ph(),data=newdf,weights=status)
summary(mod1_func)
pdf(NULL)
aa_clf<-plot(mod1_func,select = 1,ylim=c(-5,5),n=m)
invisible(dev.off())
betaest_coxlf<-as.numeric(mod1_func$coefficients[1:2])
betaests_coxlf<-as.numeric(aa_clf[[1]]$fit)
#Estimated effect from Cox Model#
plot(S,betaS,col=1,type="l")
lines(S,betaests_coxlf,col=3)

#baseline survival
#plot(mod1_func$family$data$tr,exp(-mod1_func$family$data$h),type="l",ylim=c(0,1),
#     xlab="time",ylab="survival")
est_coxlf_base<-exp(-mod1_func$family$data$h)
est_coxlf_base<-est_coxlf_base[order(mod1_func$family$data$tr)]
time_coxlf<-sort(mod1_func$family$data$tr)
#plot(time_coxlf,est_coxlf_base,type="l",col=3)
```

Next we source the file ``sourceFTTM.R" which contains all the required R functions for estimation.

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE, tidy=FALSE}
source("sourceFTTM.R")
plot(time_coxlf,est_coxlf_base,type="l",col=3)
##true baseline survival S(t)=F_epsbar(H(t)), H(t)=beta0+log(t)
true_H<-beta[1]+log(time_coxlf)
trubas<-F_epsilon_bar(true_H,r=0)
lines(time_coxlf,trubas,col="black") #matches
tau<-ceiling(max(T_obs))
p1<-ncol(Xmat)

N0grid<-c(7,10)
N1grid<-c(3,5) #can make big given resources
#r=0 #Cox specification
rgrid<-0
gridN<-expand.grid(N0=N0grid,N1=N1grid,r=rgrid)
gridN<-as.matrix(gridN)
AICres<-apply(gridN,1,ChooseAICval)  
AICvec<-c()
for (k in 1:nrow(gridN))
{AICvec[k]<-AICres[[k]]$AICval
}
indmin<-which.min(AICvec)
Nrvecsel<-gridN[indmin,] #optimal N with minimum AIC
N0<-as.numeric(Nrvecsel[1])
N1<-as.numeric(Nrvecsel[2])
r<-as.numeric(Nrvecsel[3])
  
finalest<-estFTTM.final(Nrvecsel)

psiest<-finalest$psiest
SEpar<-finalest$SEpar
VCOVpar<-finalest$VCOVpar

betaest<-psiest[1:p1] 
### estimated beta ###
betaest
#true beta#
beta[2:3]
##############
p1<-ncol(Xmat)
etaest<-psiest[(p1+1):(p1+N0+1)]
gammaest<-c()
gammaest[1]<-etaest[1]
for (k in 2: length(etaest))
{gammaest[k]<-gammaest[k-1]+exp(etaest[k])
}

thetaest<-psiest[(p1+N0+1+1):(p1+N0+1+N1+1)]
betas_tt_est<-beta_s_estfun(S,thetaest,N1)

###covariance matrix of parameters##
VCOVparfunc<-VCOVpar[(p1+N0+1+1):(p1+N0+1+N1+1),(p1+N0+1+1):(p1+N0+1+N1+1)]
basismat <- NULL
for(k in 0:N1){
  basismat<- cbind(basismat, dbeta((S), shape1=(k+1), shape2=N1-k+1))
}
basismat<-(1/(N1+1))*basismat
vcov_betas<-basismat%*%VCOVparfunc%*%t(basismat)
SE_betas<-sqrt(diag(vcov_betas))
#asymptotic 95% point-wise confidence interval
lcb_betas<-betas_tt_est-1.96*SE_betas
ucb_betas<-betas_tt_est+1.96*SE_betas
plot(S,betaS,col=1,type="l",ylim=c(-2,2)) 
lines(S,betas_tt_est,col=4)
lines(S,lcb_betas,col=4,lty=3)
lines(S,ucb_betas,col=4,lty=3) #CI looks good
abline(h=0)
```

# Example 2: Functional PO Model
Next, we generate survival times following a functional proportional odds model (Scenario A2 in paper)and fit the FTTM. 
```{r, tidy = FALSE,message=FALSE,warning=FALSE, fig.width=6, fig.height=5}
###########Generating data##############################
set.seed(1)
n<-500 
Z1<-rbinom(n,1,0.5)
Z2<-rnorm(n,0,1)
X<-cbind(Z1,Z2)
###############################
m <- 101
S<- seq(0, 1, length = m)
rankX <- 10
Phi <- cbind(1/sqrt(m), poly(S, degree=rankX-1))
lambda <- rankX:1
Xi <- sapply(lambda, function(l)
  scale(rnorm(n, sd=2*sqrt(l)), scale=FALSE))
Xs <- Xi %*% t(Phi)
beta<-c(-0.8,1.6)
beta_s<-function(s){2*sin (pi*s)} #functional effect
betaS<-beta_s(S)
lfterm2<-as.numeric((1/m)*Xs%*%betaS) 
totterm<-function(i){exp(-1*((sum(beta*X[i,])+lfterm2[i])))
}
tottermvec<-unlist(lapply(1:n,totterm)) 
mu_c<-5   
T_tr<-c() 
C<-c() 
T_obs<-c() #observed survival time
for(i in 1:n)
{
  u<-runif(1,0,1)
  temp<-sqrt((u/(1-u))*tottermvec[i] )
  T_tr[i]<- temp #true survival time
  C[i]<-rexp(1,rate=(1/mu_c))
  T_obs[i]=min(T_tr[i],C[i])
}
delta<-as.numeric(T_tr<C)
mean(delta==1) #proportion uncensored
mean(delta==0) #proportion censored 
###KM estimates#######
library(survival)
fitKM <- survfit(Surv(T_obs,delta) ~ 1)###
plot(fitKM, xlab="Days", main = 'Kaplan Meyer Plot') 
```

Next we fit functional Cox model and functional principal component based PO model for illustration. 

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE, tidy=FALSE}
newdf<-data.frame("time"=T_obs,"status"=delta,"X1"=X[,1],"X2"=X[,2])
Xmat<-X[,c(1:2)]
library(mgcv)
Smat<-matrix(S,n,m,byrow=TRUE)
Xs_sc<-Xs/m
newdf$Xs_sc<-Xs_sc
###misspecified cox model####
mod1_func <-gam(time~X1+X2+s(Smat,k=20,by=Xs_sc,bs="ps",m=2),
                family=cox.ph(),data=newdf,weights=status)
summary(mod1_func)
pdf(NULL)
aa_clf<-plot(mod1_func,select = 1,ylim=c(-5,5),n=m)
invisible(dev.off())
betaest_coxlf<-as.numeric(mod1_func$coefficients[1:2])
betaests_coxlf<-as.numeric(aa_clf[[1]]$fit)
plot(S,betaS,col=1,type="l",ylim=c(-5,5))
lines(S,betaests_coxlf,col=3) 
#PO model on FPC scores#
library(refund)
fpcamod<-fpca.face(Xs,pve=0.95) 
scoremat<-fpcamod$scores
k<-ncol(scoremat)
eigenmat<-fpcamod$efunctions
library(mets)
newdf$scoremat<-scoremat
out1 <- logitSurv(Surv(time,status)~X1+X2+scoremat,data=newdf)
summary(out1)
beta_PO<-as.numeric(coef(out1))
betaest_PO<-beta_PO[1:2]
betaest_PO_vec<-beta_PO[3:(2+k)]
betaests_PO<-m*as.numeric(eigenmat%*%betaest_PO_vec)
lines(S,betaests_PO,col=4) 
```

Next we apply the proposed FTTM for PO model specification ($r=1$).

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE, tidy=FALSE}
source("sourceFTTM.R")
tau<-ceiling(max(T_obs))
p1<-ncol(Xmat)

N0grid<-c(7,10)
N1grid<-c(3,5) #can make big given resources
#r=0 #Cox specification
rgrid<-1
gridN<-expand.grid(N0=N0grid,N1=N1grid,r=rgrid)
gridN<-as.matrix(gridN)
AICres<-apply(gridN,1,ChooseAICval)  
AICvec<-c()
for (k in 1:nrow(gridN))
{AICvec[k]<-AICres[[k]]$AICval
}
indmin<-which.min(AICvec)
Nrvecsel<-gridN[indmin,] #optimal N with minimum AIC
N0<-as.numeric(Nrvecsel[1])
N1<-as.numeric(Nrvecsel[2])
r<-as.numeric(Nrvecsel[3])
  
finalest<-estFTTM.final(Nrvecsel)

psiest<-finalest$psiest
SEpar<-finalest$SEpar
VCOVpar<-finalest$VCOVpar

betaest<-psiest[1:p1] 
### estimated beta ###
betaest
#true beta#
beta
##############
p1<-ncol(Xmat)
etaest<-psiest[(p1+1):(p1+N0+1)]
gammaest<-c()
gammaest[1]<-etaest[1]
for (k in 2: length(etaest))
{gammaest[k]<-gammaest[k-1]+exp(etaest[k])
}

thetaest<-psiest[(p1+N0+1+1):(p1+N0+1+N1+1)]
betas_tt_est<-beta_s_estfun(S,thetaest,N1)

###covariance matrix of parameters##
VCOVparfunc<-VCOVpar[(p1+N0+1+1):(p1+N0+1+N1+1),(p1+N0+1+1):(p1+N0+1+N1+1)]
basismat <- NULL
for(k in 0:N1){
  basismat<- cbind(basismat, dbeta((S), shape1=(k+1), shape2=N1-k+1))
}
basismat<-(1/(N1+1))*basismat
vcov_betas<-basismat%*%VCOVparfunc%*%t(basismat)
SE_betas<-sqrt(diag(vcov_betas))
#asymptotic 95% point-wise confidence interval
lcb_betas<-betas_tt_est-1.96*SE_betas
ucb_betas<-betas_tt_est+1.96*SE_betas
plot(S,betaS,col=1,type="l",ylim=c(-2,2)) 
lines(S,betas_tt_est,col=4)
lines(S,lcb_betas,col=4,lty=3)
lines(S,ucb_betas,col=4,lty=3) #CI looks good
abline(h=0)
```

We observe that the proposed FTTM captures the functional parameter more accurately.
<br><br><br>




