---
title: "Simulation Results"
author: "Weijia Qian"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, repr.plot.width = 12, repr.plot.height = 8)
library(CCMHr)
library(gt)
library(kableExtra)
library(MASS)
library(mgcv)
library(refund)
library(scam)
library(splines)
library(tidyverse)

folder_path <- "/Users/weijia/Research/FDA/Simulations/Output/20250716_fttm"
rda_files <- list.files(path = folder_path, pattern = "\\.RDA$", full.names = TRUE)
all_results <- list()
for (i in 1:length(rda_files)) {
  all_results[[i]] <- loadRDa(rda_files[[i]])
}
```

### Function to Create Summary Table
```{r, include=FALSE}
# Transpose dataset from wide to long
create_table <- function(data, var, beta_type, title){
  data1 <- data %>%
  pivot_longer(cols = var,
               names_to = "type",
               values_to = "value") %>%
  mutate(est_model = case_when(str_detect(type, "cox") ~ "mgcv LFCM",
                               str_detect(type, "fttm") ~ "FTTM",
                               str_detect(type, "norm") ~ "mgcv lfAFT",
                               str_detect(type, "faft2") ~ "loglogistic lfAFT",
                               .default = "lognormal lfAFT"))

  # Combine 'n' and 'nS' into wider columns
  data2 <- data1 %>%
    #filter(beta_type == !!beta_type) %>%
    pivot_wider(names_from = c(n, nS, beta_type), values_from = value, names_sep = "_nS_")

  gt_table <- data2 %>%
    dplyr::select(-c(type)) %>%
    gt(groupname_col = "family") %>%
    tab_spanner(label = "N = 100", columns = starts_with("100_")) %>%
    tab_spanner(label = "N = 200", columns = starts_with("200_")) %>%
    tab_spanner(label = "N = 500", columns = starts_with("500_")) %>%
    cols_label(ends_with("nS_50") ~ "nS = 50",
               ends_with("nS_100") ~ "nS = 100",
               ends_with("nS_500") ~ "nS = 500") %>%
   tab_header(title = title) %>%
   cols_align(align = "center", columns = everything()) %>%
   fmt_number(columns = where(is.numeric), decimals = 3) %>%
   opt_table_outline() %>%
   opt_row_striping()

  return(gt_table)
}
```


### Average Out-of-Sample Brier Score Across Iterations

```{r}
all_info <- data.frame()
# Loop through all scenarios and iterations to extract 'info'
for (scenario in all_results) {
  print(scenario[[1]]$info$scenario)
  for (iteration in scenario) {
    all_info <- rbind(all_info, iteration$info)
  }
}

# Calculate the average Brier Score Across iterations
table_Brier <- all_info %>%
  filter(if_all(where(is.numeric), ~ !is.infinite(.))) %>%
  group_by(beta_type, family, n, nS) %>%
  summarise(
    Brier_norm = round(mean(Brier_norm, na.rm = TRUE), 3),
    Brier_cox = round(mean(Brier_cox, na.rm = TRUE), 3),
    Brier_faft = round(mean(Brier_faft, na.rm = TRUE), 3),
    Brier_faft2 = round(mean(Brier_faft2, na.rm = TRUE), 3),
    Brier_fttm = round(mean(Brier_fttm, na.rm = TRUE), 3),
    .groups = "drop"
  )

# Create summary table
# print(table_Brier)

# plot
plot_Brier <- table_Brier %>%
  pivot_longer(cols = c("Brier_norm", "Brier_cox", "Brier_faft", "Brier_faft2", "Brier_fttm"),
               names_to = "type",
               values_to = "value") %>%
  mutate(est_model = case_when(str_detect(type, "cox") ~ "mgcv LFCM",
                               str_detect(type, "fttm") ~ "FTTM",
                               str_detect(type, "norm") ~ "mgcv lfAFT",
                               str_detect(type, "faft2") ~ "loglogistic lfAFT",
                               .default = "lognormal lfAFT"))

plot_Brier %>%
  ggplot(aes(factor(n), value, fill = est_model)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_grid(family~beta_type, labeller = label_both) +
  labs(x = "n", y = "Brier Score", title = "Average Brier Score Across 100 Simulations", fill = "Model")

```

### Plot Estimated Coefficient + CI + True Coefficient
Fitted using the log-normal functional AFT model.

```{r}
sub_coef <- data.frame()

# Loop through all scenarios and iterations to extract 'coef'
for (scenario in all_results) {
  counter <- 0  # Initialize a counter

  for (iteration in scenario) {
    # Break the inner loop if 50 iterations for the current scenario are stored
    if (counter >= 1) {
      break
    }

    # Extract data for the current iteration
    time <- seq(0, 1, length.out = iteration$info$nS)
    tmp_data <- data.frame(scenario = iteration$info$scenario,
                           iter = iteration$info$iter,
                           family = iteration$info$family,
                           beta_type = iteration$info$beta_type,
                           n = iteration$info$n,
                           nS = iteration$info$nS,
                           b = iteration$info$b,
                           time = time,
                           censor_rate = iteration$info$censor_rate,
                           true_coef = iteration$coef$true_coef,
                           est_coef_norm = iteration$coef$est_coef_norm,
                           lb_coef_norm = iteration$coef$lb_coef_norm,
                           ub_coef_norm = iteration$coef$ub_coef_norm,
                           est_coef_faft = iteration$coef$est_coef_faft,
                           lb_coef_faft = iteration$coef$lb_coef_faft,
                           ub_coef_faft = iteration$coef$ub_coef_faft,
                           est_coef_faft2 = iteration$coef$est_coef_faft2,
                           lb_coef_faft2 = iteration$coef$lb_coef_faft2,
                           ub_coef_faft2 = iteration$coef$ub_coef_faft2,
                           est_coef_fttm = iteration$coef$est_coef_fttm,
                           lb_coef_fttm = iteration$coef$lb_coef_fttm,
                           ub_coef_fttm = iteration$coef$ub_coef_fttm,
                           cma_lb_coef_norm = iteration$coef$cma_lb_coef_norm,
                           cma_ub_coef_norm = iteration$coef$cma_ub_coef_norm,
                           est_coef_cox = iteration$coef$est_coef_cox,
                           lb_coef_cox = iteration$coef$lb_coef_cox,
                           ub_coef_cox = iteration$coef$ub_coef_cox,
                           cma_lb_coef_cox = iteration$coef$cma_lb_coef_cox,
                           cma_ub_coef_cox = iteration$coef$cma_ub_coef_cox)

    # Append to all_coef
    sub_coef <- rbind(sub_coef, tmp_data)

    # Increment the counter
    counter <- counter + 1
  }
}

# Ensure the required directory for saving plots exists
# output_dir <- "../Figures"
# if (!dir.exists(output_dir)) {
#   dir.create(output_dir)
# }

# select a subset of scenarios
# sub_coef_2 <- sub_coef %>%
#   filter(family != "cox.ph")

# Loop through each unique scenario
scenarios <- unique(sub_coef[, c("family", "n", "nS", "beta_type", "b", "censor_rate")])
library(scales)
for (i in 1:nrow(scenarios)) {
  # Extract the scenario information
  scenario <- scenarios[i, ]
  family <- scenario$family
  n <- scenario$n
  nS <- scenario$nS
  beta_type <- scenario$beta_type
  b <- scenario$b
  censor_rate <- scenario$censor_rate

# Filter data for the current scenario
  scenario_data <- sub_coef %>%
    filter(family == !!family, n == !!n, nS == !!nS, beta_type == !!beta_type, b == !!b)

# Create the plot
p <- ggplot(data = scenario_data, aes(x = time)) +
  #geom_ribbon(aes(ymin = lb_coef_faft, ymax = ub_coef_faft, fill = "Pointwise CI"), alpha = 0.5) +
  #geom_ribbon(aes(ymin = cma_lb_coef_norm, ymax = cma_ub_coef_norm, fill = "CMA CI"), alpha = 0.3) +
  geom_line(aes(y = true_coef), color = "red", linetype = "dashed", size = 1) +
  geom_line(aes(y = est_coef_norm, color = "mgcv lfAFT"), size = 0.7) +
  geom_line(aes(y = est_coef_cox, color = "mgcv LFCM"), size = 0.7) +
  geom_line(aes(y = est_coef_faft, color = "lognormal lfAFT"), size = 0.7) +
  geom_line(aes(y = est_coef_faft2, color = "loglogistic lfAFT"), size = 0.7) +
  geom_line(aes(y = est_coef_fttm, color = "FTTM"), size = 0.7) +
  scale_color_manual(values = hue_pal()(6)[-1]) +
  # scale_fill_manual(values = c(
  #   "Pointwise CI" = "lightblue"
  #   #"CMA CI" = "lightgreen"
  # )) +
  labs(
    x = "Time",
    y = "Coefficient",
    title = paste("family =",family, "n =", n, "nS =", nS, "beta_type =", beta_type),
    subtitle = paste("b=", b, "censor rate =", censor_rate),
    color = NULL,
    fill = NULL
  ) +
  theme_minimal()
print(p)
}

# Save the plot
# file_name <- paste0(output_dir, "/", family, "_", n, "_", nS, "_", beta_type, ".png")
# ggsave(file_name, plot = p, width = 8, height = 6)
```



### MISE for Estimated Coefficient Functions Across Iterations
```{r}
all_coef <- data.frame() 

# Loop through all scenarios and iterations to extract 'coef'
for (scenario in all_results) {
  for (iteration in scenario) {
    
    # Extract data for the current iteration
    tmp_data <- data.frame(scenario = iteration$info$scenario,
                           iter = iteration$info$iter,
                           family = iteration$info$family,
                           n = iteration$info$n,
                           nS = iteration$info$nS,
                           b = iteration$info$b,
                           #time = time,
                           beta_type = iteration$info$beta_type,
                           mise_coef_norm = mean(iteration$coef$se_coef_norm),
                           mise_coef_cox = mean(iteration$coef$se_coef_cox),
                           mise_coef_faft = mean(iteration$coef$se_coef_faft),
                           mise_coef_faft2 = mean(iteration$coef$se_coef_faft2),
                           mise_coef_fttm = mean(iteration$coef$se_coef_fttm),
                           cp_coef_norm = mean(iteration$coef$cover_coef_norm),
                           cp_coef_cox = mean(iteration$coef$cover_coef_cox),
                           cp_coef_faft = mean(iteration$coef$cover_coef_faft),
                           cp_coef_faft2 = mean(iteration$coef$cover_coef_faft2),
                           cp_coef_fttm = mean(iteration$coef$cover_coef_fttm),
                           cp_cma_coef_norm = all(iteration$coef$cover_cma_coef_norm),
                           cp_cma_coef_cox = all(iteration$coef$cover_cma_coef_cox))

    # Append to all_coef
    all_coef <- rbind(all_coef, tmp_data)
  }
}

# Calculate the average MISE of coefficient function Across iterations
table_mise_coef <- all_coef %>%
  group_by(b, beta_type, family, n, nS) %>%
  summarise(
    mise_coef_norm = round(mean(mise_coef_norm, na.rm = TRUE), 3),
    mise_coef_cox = round(mean(mise_coef_cox, na.rm = TRUE), 3),
    mise_coef_faft = round(mean(mise_coef_faft, na.rm = TRUE), 3),
    mise_coef_faft2 = round(mean(mise_coef_faft2, na.rm = TRUE), 3),
    mise_coef_fttm = round(mean(mise_coef_fttm, na.rm = TRUE), 3),
    .groups = "drop"
  ) 

# plot
plot_mise_coef <- table_mise_coef %>%
  pivot_longer(cols = c("mise_coef_norm", "mise_coef_cox", "mise_coef_faft", "mise_coef_faft2", "mise_coef_fttm"),
               names_to = "type",
               values_to = "value") %>%
  mutate(est_model = case_when(str_detect(type, "cox") ~ "mgcv LFCM",
                               str_detect(type, "fttm") ~ "FTTM",
                               str_detect(type, "norm") ~ "mgcv lfAFT",
                               str_detect(type, "faft2") ~ "loglogistic lfAFT",
                               .default = "lognormal lfAFT"))

plot_mise_coef %>%
  ggplot(aes(factor(n), value, fill = est_model)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_grid(family~beta_type, labeller = label_both) +
  labs(x = "n", y = "MISE", title = "Average MISE of Coefficient Function Across 100 Simulations", fill = "Model")
```

### Average Coverage Probability of Pointwise CIs for Coefficient Function
```{r}
# Calculate the average CP of Pointwise CIs for Coefficient Function
table_cp_coef <- all_coef %>%
  group_by(b, beta_type, family, n, nS) %>%
  summarise(
    cp_coef_norm = round(mean(cp_coef_norm, na.rm = TRUE), 3),
    cp_coef_cox = round(mean(cp_coef_cox, na.rm = TRUE), 3),
    cp_coef_faft = round(mean(cp_coef_faft, na.rm = TRUE), 3),
    cp_coef_faft2 = round(mean(cp_coef_faft2, na.rm = TRUE), 3),
    cp_coef_fttm = round(mean(cp_coef_fttm, na.rm = TRUE), 3),
    .groups = "drop"
  ) 

# plot
plot_cp_coef <- table_cp_coef %>%
  pivot_longer(cols = c("cp_coef_norm", "cp_coef_cox", "cp_coef_faft", "cp_coef_faft2", "cp_coef_fttm"),
               names_to = "type",
               values_to = "value") %>%
  mutate(est_model = case_when(str_detect(type, "cox") ~ "mgcv LFCM",
                               str_detect(type, "fttm") ~ "FTTM",
                               str_detect(type, "norm") ~ "mgcv lfAFT",
                               str_detect(type, "faft2") ~ "loglogistic lfAFT",
                               .default = "lognormal lfAFT"))

plot_cp_coef %>%
  ggplot(aes(factor(n), value, fill = est_model)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_hline(yintercept=0.95, linetype="dashed", color = "red") +
  facet_grid(family~beta_type, labeller = label_both) +
  labs(x = "n", y = "Coverage Probability", title = "Average Pointwise Coverage Probability of Coefficient Function Across 100 Simulations", fill = "Model")
```

### MISE for Estimated Survival Functions
```{r}
all_surv <- data.frame() 

# Loop through all scenarios and iterations to extract 'coef'
for (scenario in all_results) {
  for (iteration in scenario) {
    # Extract data for the current iteration
    tmp_data <- data.frame(scenario = iteration$info$scenario,
                           iter = iteration$info$iter,
                           family = iteration$info$family,
                           n = iteration$info$n,
                           nS = iteration$info$nS,
                           beta_type = iteration$info$beta_type,
                           b = iteration$info$b,
                           mise_surv_norm = mean(iteration$surv$se_surv_norm),
                           mise_surv_cox = mean(iteration$surv$se_surv_cox),
                           mise_surv_faft = mean(iteration$surv$se_surv_faft),
                           mise_surv_faft2 = mean(iteration$surv$se_surv_faft2),
                           mise_surv_fttm = mean(iteration$surv$se_surv_fttm))

    # Append to all_coef
    all_surv <- rbind(all_surv, tmp_data)
  }
}

table_mise_surv <- all_surv %>%
  group_by(b, beta_type, family, n, nS) %>%
  summarise(
    mise_surv_norm = round(mean(mise_surv_norm, na.rm = TRUE), 3),
    mise_surv_cox = round(mean(mise_surv_cox, na.rm = TRUE), 3),
    mise_surv_faft = round(mean(mise_surv_faft, na.rm = TRUE), 3),
    mise_surv_faft2 = round(mean(mise_surv_faft2, na.rm = TRUE), 3),
    mise_surv_fttm = round(mean(mise_surv_fttm, na.rm = TRUE), 3),
    .groups = "drop"
  )

# plot
plot_mise_surv <- table_mise_surv %>%
  pivot_longer(cols = c("mise_surv_norm", "mise_surv_cox", "mise_surv_faft", "mise_surv_faft2", "mise_surv_fttm"),
               names_to = "type",
               values_to = "value") %>%
  mutate(est_model = case_when(str_detect(type, "cox") ~ "mgcv LFCM",
                               str_detect(type, "fttm") ~ "FTTM",
                               str_detect(type, "norm") ~ "mgcv lfAFT",
                               str_detect(type, "faft2") ~ "loglogistic lfAFT",
                               .default = "lognormal lfAFT"))

plot_mise_surv %>%
  ggplot(aes(factor(n), value, fill = est_model)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_grid(family~beta_type, labeller = label_both) +
  labs(x = "n", y = "MISE", title = "Average MISE of Survival Function Across 100 Simulations", fill = "Model")
```



